---
title: "Practical Machine Learning: Prediction Model Assignment Write-up"
author: "Ron Rick Chua"
date: "Monday, August 18, 2014"

---


## Synopsis:

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset). 


We will use the data mentioned above to develop a predictive model for this classes.



Loading library used.
```{r}
library(ggplot2)
library(lattice)
library(caret)
library(randomForest)
library(rpart)
```
Clearing Workspace
```{r}
rm(list=ls())
```
Loading the Data Sets of Testing and Training to R. 
```{r}
TrainingSet <-read.csv("pml-training.csv")
TestingSet  <-read.csv("pml-testing.csv" )
```



## Exploring the Data Set
For Training Set:
```{r}
# Number of rows:
nrow(TrainingSet)

# Number of columns:
ncol(TrainingSet)

#Summary details:
summary(TrainingSet[,c(1:2,159:160)])
```
Using a Scatterplot Matrices for visual reference
```{r}
pairs(TrainingSet[1:10000,1:10])
pairs(TrainingSet[1:10000,11:20])
pairs(TrainingSet[1:10000,21:30])
```

## Getting and Cleaning the Data Set
There are 160 columns in the Training Set, and we will be excluding columns with NA values.
```{r}
NewTraining<-TrainingSet
NewTraining[ NewTraining == '' | NewTraining == 'NA'] <- NA
indx <-which(colSums(is.na(NewTraining))!=0)
NewTraining<-NewTraining[, -indx]
NewTraining<-NewTraining[,-(1:7)]

```
## Validate Data Set
Creating a Data set that are Valid, This is for Cross validating with the Training Set.
```{r}
InTraining  <- createDataPartition(y=NewTraining$classe,p=0.70,list=FALSE)
NewTraining <- NewTraining[InTraining,]
ValidateSet <- NewTraining[-InTraining,]
```

## Prediction Model Set
Random Forest we will be using to train our Prediction Model set to predict the weight lifting quality in the Training Set.
```{r}
Pmodel <- train(classe~., data=NewTraining, method = "rf", tuneLength = 1, ntree = 25)
print(Pmodel)
```

## Evaluating the Model
Using Confussion Matrix to evaluate the Prediction Model set versus the Validate Data set.
```{r}
confusionMatrix(predict(Pmodel, ValidateSet), ValidateSet$classe)
```

Visual Representation of accuracy


```{r}
plot(predict(Pmodel,newdata=ValidateSet[,-ncol(ValidateSet)]),ValidateSet$classe, xlab="Validate Set", ylab="Prediction Model Set",col = c("black","red", "blue","green","yellow"))
```
```
Note on Class:
A: Exactly according to the specification
B: Throwing the elbows to the front
C: Lifting the dumbbell only halfway
D: Lowering the dumbbell only halfway
E: Throwing the hips to the front
```

## Compuring for the Accuracy of the Model
```{r}
accurate <- c(as.numeric(predict(Pmodel,newdata=ValidateSet[,-ncol(ValidateSet)])==ValidateSet$classe))
MAccuracy <- sum(accurate)*100/nrow(ValidateSet)
message("Accuracy of Prediction Model set VS Validate Data set = ", format(round(MAccuracy, 2), nsmall=2),"%")
```
# Conclusion
A 100% accuracy was computed here,  but caution must be taken due to the use of Random forest, tends to Overfitting.


```


```



# Predictions on the testing set:

For Testing Set:
```{r}
# Number of rows:
nrow(TestingSet)

# Number of columns:
ncol(TestingSet)

# Summary details:
summary(TestingSet[,c(1:2,159:160)])
```
## Results of the Prediction to Testing Set
```{r}
Ptest<-predict(Pmodel, TestingSet)
print(Ptest)
```


```


```

# References


The Data used for this project can be downloaded at:

Training set
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv


Test Set
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv


I would like to cite Human Activity Recognition publication and the Major Collaborators:

- Wallace Ugulino 
- Eduardo Velloso
- Hugo Fuks

Which allows us to use their data set for our course project.

Web Site: http://groupware.les.inf.puc-rio.br/har#ixzz3AR2M0igh


